Echoes from the scattering of wavepackets on wormholes

### 2019-01-19 paper2.tex
I think it is now OK...

The layout for this paper is as follows: in section \ref{sec:scalar}, we review the scattering problem of scalar waves starting by a quick overview of the dispersion of a Gaussian pulse by a Schwarzschild black hole. The main point of this section is to calculate the transmission and reflection coefficients of each of the centrifugal barriers constituting the resonant cavity, formed in the case of a wormhole. Our results show a frequency ``sweet spot'' such that the incident pulse is not fully reflected nor fully transmitted by the cavity, favouring multiple internal reflections that source the QNMs.  Furthermore, we solve the scattering problem by a wormhole directly using the same ingoing Gaussian wavepacket, and then we compare the Fourier transform of this solution with the pulse reconstructed following the geometrical optics approximation. We find that the approximate reconstruction matches the full solution, up to the peaks due to the QNM frequencies. Likewise, we evaluate the amplitude of each of the echoes as a function of the width of the initial gaussian waveform, finding that a single width of the incident pulse maximizes the amplitude of each echo. In section \ref{sec:tensor}, we extend the previous results in the previous section for a Gaussian pulse of tensor fluctuations of the metric by following the even and odd decomposition of the tensor modes introduced by Regge, Wheeler and Zerilli in \citep{Regge:1957td, PhysRevD.2.2141, PhysRevD.5.2419, PhysRevD.5.2439}. Our results can be recasted in terms of the usual asymptotic polarization modes $h_+$ and $h_{\times}$, known as the perturbations of a flat metric. Finally, in section \ref{sec:conclusions}, we discuss and conclude.

@@

Amplitudes of the outgoing signals depend on the width of the incident Gaussian wavepacket, this is visible in Fig.~\ref{fig:asympt_two_sigmas} where we plot the asymptotic solutions for two different ingoing wavepackets: one with $\sigma=0.6495r_g$ in the left panel and a second one with $\sigma=5.196r_g$ in the right. After subtracting the outgoing solution for a black hole, the presence of a train of wavelets, widely known as echoes, is very clear. For large values of $\sigma$, the signal is not featureless after the transient, as we can observe in the right panel of the same figure. Therefore, subtracting the black hole outgoing pulse (i.e., the case in which there is only one potential wall) turns to be a convenient way to clean the signal from back reflections due to the ``tails'' of the potential barrier, especially since in this case the echoes are four orders of magnitude smaller than the transient. In both panels, we plot the variable $\Delta\mathcal{U}_{20}|_{\mathrm{clean}}\equiv\mathcal{U}_{20}|_{\mathrm{original}}-\mathcal{U}^{\mathrm{bh}}_{20}$ in the right upper corner of the figures to represent the echoes and their net amplitude after removing backscattering effects. Notice that, in the upper corner of both figures, the amplitude of the echoes does not decrease exponentially with time due to the successive reflections inside the cavity.

@@

In analogy with the previous section, now we solve the equations of motion for the scattering process. Our setup for the initial conditions of $\Psi_{20}^{\mathrm{odd}}(x,0)$ and $\Psi_{20}^{\mathrm{even}}(x,0)$ and their time derivatives is not different from \eqref{eq:scalar_init_cond}  
\begin{align}
\displaystyle{\Psi^{\mathrm{odd}}_{20}(x,0)=\exp\left(\frac{(x-x_0)^2}{2\sigma^2}\right)},\nonumber\\
\displaystyle{\partial_t \Psi^{\mathrm{odd}}_{20}(x,t)\bigg{|}_{t=0}=\partial_x\Psi^{\mathrm{odd}}_{20}(x,0)},
\label{eq:tensor_init_cond}
\end{align}
and the same applies for $\Psi^\mathrm{even}_{20}$ and its initial time derivative. Using $\sigma=0.9185r_g$, the same initial position of the Gaussian wavepackets -- i.e., $x_0=60.0r_g$ -- and the same separation between potential walls -- i.e., $r_0=20.0r_g$ -- as before. We show the evolution of $\Psi_{20}^{\mathrm{even}}(x,t)$ in Fig.~\ref{fig:even_evolve}, where the dispersion of the ingoing pulse is not significantly different from our results in the left panel of Fig.~\ref{fig:sol_n_flux_wh}: this is not surprising due to the similarities between the shapes of the effective potentials for scalar and tensor modes, which seem to become even more similar for higher values of $\ell$. At late times, the cavity is filled showing an interference pattern. Internal reflections make the QNMs propagate for longer in the spheres of maximum effective potential.

>>>>

The layout for this paper is as follows: in section \ref{sec:scalar}, we review the scattering problem of scalar waves starting by a quick overview of the dispersion of a Gaussian pulse by a Schwarzschild black hole. The main point of this section is to calculate the transmission and reflection coefficients of each of the centrifugal barriers constituting the resonant cavity, formed in the case of a wormhole. Our results show a frequency ``sweet spot'' such that the incident pulse is not fully reflected nor fully transmitted by the cavity, favouring multiple internal reflections that source the QNMs.  Furthermore, we solve the scattering problem by a wormhole directly using the same ingoing Gaussian wavepacket, and then we compare the Fourier transform of this solution with the pulse reconstructed following the geometrical optics approximation. We find that the approximate reconstruction matches the full solution, up to the peaks due to the QNM frequencies. Likewise, we evaluate the amplitude of each of the echoes as a function of the width of the initial gaussian waveform, finding that a single width of the incident pulse maximizes the amplitude of all the echoes. In section \ref{sec:tensor}, we extend the previous results in the previous section for a Gaussian pulse of tensor fluctuations of the metric by following the even and odd decomposition of the tensor modes introduced by Regge, Wheeler and Zerilli in \citep{Regge:1957td, PhysRevD.2.2141, PhysRevD.5.2419, PhysRevD.5.2439}. Our results can be recasted in terms of the usual asymptotic polarization modes $h_+$ and $h_{\times}$, known as the perturbations of a flat metric. Finally, in section \ref{sec:conclusions}, we discuss and conclude.

@@

Amplitudes of the outgoing signals depend on the width of the incident Gaussian wavepacket, this is visible in Fig.~\ref{fig:asympt_two_sigmas} where we plot the asymptotic solutions for two different ingoing wavepackets: one with $\sigma=0.6495r_g$ in the left panel and a second one with $\sigma=5.196r_g$ in the right. After subtracting the outgoing solution for a black hole, the presence of a train of wavelets, colloquially known as echoes, is very clear. For large values of $\sigma$, the signal is not featureless after the transient, as we can observe in the right panel of the same figure. Therefore, subtracting the outgoing pulse (i.e., the case in which there is only one potential wall) obtained from the black hole is a convenient way to clean the signal from back reflections due to the ``tails'' of the potential barrier. In the particular case depicted in the right panel of Fig.~\ref{fig:asympt_two_sigmas}, the amplitudes of the echoes are four orders of magnitude smaller than the transient. In both panels, we plot the variable $\Delta\mathcal{U}_{20}|_{\mathrm{clean}}\equiv\mathcal{U}_{20}|_{\mathrm{original}}-\mathcal{U}^{\mathrm{bh}}_{20}$ in the right upper corner of the figures to represent the echoes and their net amplitude after removing backscattering effects. Notice that, in the upper corner of both figures, the amplitude of the echoes does not decrease exponentially with time due to the successive reflections inside the cavity.

@@

In analogy with the previous section, now we solve the equations of motion for the scattering process. Our setup for the initial conditions of $\Psi_{20}^{\mathrm{odd}}(x,0)$ and $\Psi_{20}^{\mathrm{even}}(x,0)$ and their time derivatives is not different from \eqref{eq:scalar_init_cond}  
\begin{align}
\displaystyle{\Psi^{\mathrm{odd}}_{20}(x,0)=\exp\left(\frac{(x-x_0)^2}{2\sigma^2}\right)},\nonumber\\
\displaystyle{\partial_t \Psi^{\mathrm{odd}}_{20}(x,t)\bigg{|}_{t=0}=\partial_x\Psi^{\mathrm{odd}}_{20}(x,0)},
\label{eq:tensor_init_cond}
\end{align}
and the same applies for $\Psi^\mathrm{even}_{20}$ and its initial time derivative. Using $\sigma=0.9185r_g$, the same initial position of the Gaussian wavepackets -- i.e., $x_0=60.0r_g$ -- and the same separation between potential walls -- i.e., $r_0=20.0r_g$ -- as before. We show the evolution of $\Psi_{20}^{\mathrm{odd}}(x,t)$ in Fig.~\ref{fig:even_evolve}, where the dispersion of the ingoing pulse is not significantly different from our results in the left panel of Fig.~\ref{fig:sol_n_flux_wh}: this is not surprising due to the similarities between the shapes of the effective potentials for scalar and tensor modes, which seem to become even more similar for higher values of $\ell$. At late times, the cavity is filled showing an interference pattern. Internal reflections make the QNMs propagate for longer in the spheres of maximum effective potential.
